---
title: "model_forcing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{model_forcing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# 1. Illustration of Meteorological forcing

```{r setup}
library(PMLV2)
library(terra)

dir_root = "PMLV2_forcing_NorthChina/MCD12Q1_006_IGBP/"

indir <- system.file("PMLV2_forcing_NorthChina/MCD12Q1_006_IGBP/", package = "PMLV2")
fs_lc <- dir(indir, "*.tif", full.names = TRUE)

indir <- system.file("PMLV2_forcing_NorthChina/mete_GLDASV21/", package = "PMLV2")
fs_mete <- dir(indir, "*.nc", full.names = TRUE)

print(basename(fs_lc))
print(basename(fs_mete))
```

Have a look about those data: 

## 1.1. IGBP

IGBP is from `MCD12Q1_006_IGBP`(<https://developers.google.com/earth-engine/datasets/catalog/MODIS_006_MCD12Q1>).

<https://code.earthengine.google.com/5fb1bd3c6245799c85e993d687dfffb2>

> 0.25°×0.25°, 18 kinds of IGBP, the fraction of each IGBP in a 0.25°×0.25° grid.
> I prepared this data on the GEE.

```{r}
r = rast(fs_lc[1])
print(r)
names(r)
```

## 1.2. Meteorological forcing

`PMLV2_forcing_NorthChina_GLDASV21_2000.nc` is a ncfile, with the variables of:

`c("LAI", "Emiss", "Albedo", "Tmax", "Tmin", "Tavg", "q", "Rln", "Rs", "Pa", "U2", "Prcp")`

> I prepared this data also on the GEE.

```{r}
r = rast(fs_mete[1])
print(r)
varnames(r)
```

### 1.2.1. Unit of input forcing

**Table 1.** Variables in the PMLV2 model.

| variable | Longname                         | Unit  |
| -------- | -------------------------------- | ----- |
| `LAI`    | Leaf are index                   | -     |
| `Emiss`  | Emissivity                       | -     |
| `Albedo` | Albedo                           | -     |
| `Tmax`   | Maximum temperature              | ℃     |
| `Tmin`   | Minimum temperature              | ℃     |
| `Tavg`   | Average temperature              | ℃     |
| `q`      | Specific humidity                | kg/kg |
| `Rln`    | Downward longwave radiation      | W m-2 |
| `Rs`     | Downward shortwave radiation     | W m-2 |
| `Pa`     | Air Pressure                     | kPa   |
| `U2`     | Wind speed in **<u>2 meter</u>** | m/s   |
| `Prcp`   | Precipitation                    | mm/d  |


## 1.3. Prepare input

### 1.3.1. Convert original data into 8-day on the GEE

1. `ERA5L hourly to 8-day`: <https://code.earthengine.google.com/693b6ee233f8cf5dcc7c2e1579631967>

2. `GLDAS 3hourly to 80day`: <https://code.earthengine.google.com/57102f28df2690cdae5b742f97561136>

> You don't have the access to GEE gee_PML2. It will not available until my new paper published.

### 1.3.2. Download them from GEE

```{r}
# go to GEE

```

### 1.3.3. Tidy downloaded tiff file, and convert them to nc
